---
title: "Global Diversity of Mesopelagic Mesozooplankton"
output: html_document
date: '2022-07-15'
---

```{r lib, message=FALSE, warning=FALSE, paged.print=FALSE}
library(readr)
library(dplyr)
library(glue)
library(latex2exp)
library(sp)
library(scales)
library(tidyr)
library(tidyverse)
library(data.table)
library(biscale)
library(cowplot)
library(ggpubr)
library(janitor)
library(sf)
library(patchwork)

# custom functions
`%notin%` <- Negate(`%in%`)
select=dplyr::select
rename=dplyr::rename
filter=dplyr::filter
summarize=dplyr::summarize
count=dplyr::count
drop_na=tidyr::drop_na
group_by=dplyr::group_by
mutate=dplyr::mutate
```

```{r functions}
# load functions to create maps and plots
#nice theme for plotting
source(here::here("fun/theme_Publication.R"))
# plot map in Eckert IV global equal area projection
source(here::here("fun/plot_map.R")) 
#plot the LDG
source(here::here("fun/ltg.R"))
# plot map in Eckert IV global equal area projection with 2D color legend
source(here::here("fun/plot.map2D.R")) 
```
```{r include=FALSE}
# from https://www.r-bloggers.com/2016/04/map-of-the-windows-fonts-registered-with-r/
# to load fonts:
# extrafont::font_import()
# Reads the fonttable database and registers those fonts with R
extrafont::loadfonts(device = "all")
```

# 1. Load Data
```{r data}
# load grid with names of mesopelagic provinces
GEO <- read_csv("data/GEO.csv",show_col_types=FALSE)
grid=GEO[, c("longitude","latitude","ProvId")]
# load meta file for zooplankton species
META_zoop <- read_csv("data/META_zoop.csv",show_col_types=FALSE)
# files species that were run in sDM and were in mesopelagic zone 
files <- META_zoop |>  filter(ModelStatus =="SDM",  Env %notin% c("EPI", "NEITHER")) |> pull(SPID)
# __ give the PORJ.4 string for Eckert IV projection
PROJ <- "+proj=eck4 +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs" 
#calculate N per phylum
n_phylum <- META_zoop |>  
  filter(ModelStatus == "SDM", Env %in% c("MESO", "BOTH")) |> 
  tabyl(phylum) |> 
  mutate(phylum_label = glue::glue("{phylum} (n={n})")) |> 
  select(phylum, phylum_label) |>
  rbind(data.frame(
    phylum = "All species",
    phylum_label = glue::glue("All species (n={length(files)})")
  ))

#load occurrence data for each species and total
ENSEMBLE_occ.wmean <- read_csv("data/ENSEMBLE_occ.wmean.csv")
ENSEMBLE_occ.wmean.tot<- read_csv("data/ENSEMBLE_occ.wmean.tot.csv")

# calculate max number of species per latitude
max.sp_lat <- ENSEMBLE_occ.wmean |> 
  pivot_longer(-c("longitude","latitude", "ProvId")) |> 
  filter(value>=1) |> 
  group_by(latitude) |>
  summarise(count = n_distinct(name),.groups = "keep")

# Sampling effort
effort.infer<- read_csv( "data/sampling-effort.csv",show_col_types = FALSE) |>   rename(n_samples_obs=total, num_sp_obs=total_occ1) |>  
  left_join(ENSEMBLE_occ.wmean.tot)  |> 
  rename(num_sp_pred=total) |> 
  mutate(effort=num_sp_obs/num_sp_pred)
```

```{r ENV}
# Load Mesopelagic Data
load(here::here("rds/Temp.RData"))
Temp
#Std Dev Temp
load("rds/Temp_mean_sd.RData")
MESO
```

### Unfiltered Data

```{r}
## prepare data to calculate the rarity
sp.importance <- ENSEMBLE_occ.wmean |>
  group_by(longitude, latitude, ProvId) |> 
  pivot_longer(-c("longitude","latitude","ProvId")) |> 
  filter(value>0)|> 
  group_by(name) |> 
  # number of cells for each species
  mutate(N=n()) |> 
  ungroup() |> 
  # proportion of distribution of each sp found in a given cell
  mutate(R=value/N) # equivalent to 1/N

# calculate the total rarity
dt<- sp.importance |> 
  group_by(latitude, longitude, ProvId) |>
  # weighted endemism
  summarize(sum = sum(R, na.rm=TRUE), .groups="keep") |> 
  ungroup() |> 
  right_join(grid, by = c("latitude", "longitude","ProvId")) |> 
  drop_na(ProvId) |> 
  select(longitude, latitude, sum)|> 
  mutate(sum=scales::rescale(sum))|> 
  drop_na()

# calculate the average rarity
dt1 <- sp.importance |> group_by(latitude, longitude, ProvId) |>
  summarize(sum = sum(R, na.rm = TRUE) / n(),
            .groups = "keep") |> 
  ungroup() |> 
  right_join(grid, by = c("latitude", "longitude", "ProvId")) |> 
  drop_na(ProvId) |>
  select(longitude, latitude, sum) |> 
  mutate(sum = scales::rescale(sum)) |> 
  drop_na()

# make a data frame that gets rarity for each lat/lon and also create a column for each ocean basin
diver_biom<- ENSEMBLE_occ.wmean.tot |> 
  rename(SR =total) |> # number of species per lat/lon
  left_join(grid|>  select("longitude","latitude","ProvId"))|>
  left_join(dt|>  rename(tot_rar=sum))|>
  left_join(dt1|>  rename(avrg_rar=sum))|> 
  mutate(
    basin = case_when(
      ProvId %in% c(1) ~ "Arctic",
      ProvId %in% c(2, 3, 4, 5, 6, 11, 12) ~ "North Pacific",
      ProvId %in% c(7, 8, 9 , 10, 13) ~ "South Pacific",
      ProvId %in% c(14, 15, 16, 17, 18, 19, 20) ~ "Indian Ocean",
      ProvId %in% c(21, 22, 23, 24, 25, 26) ~ "North Atlantic",
      ProvId %in% c(27, 28, 29, 30) ~ "South Atlantic",
      ProvId %in% c(31, 32, 33) ~ "Southern Ocean",
    )) 

diver_biom
```

### Filtered Data

Here and after filtered data just remove two species from the maps: Allosergestes pestafer and Echinomysis serratus. All steps are same as in Unfilttered sections (see explanations for each step there)
```{r}
# Filtered data : removes 2 species from S. Pacific
sp.importance.fil <- ENSEMBLE_occ.wmean |>
  group_by(longitude, latitude, ProvId) |> 
  pivot_longer(-c("longitude","latitude","ProvId")) |> filter(value>0)|> 
  filter(name %notin% c("Allosergestes.pestafer_wmean", "Echinomysis.serratus_wmean")) |>
  group_by(name) |> 
  # number of cells for each species
  mutate(N=n()) |> 
  ungroup() |> 
  # proportion of distribution of each sp found in a given cell
  mutate(R=value/N) # equivalent to 1/N

# calculate the total rarity
dt.fil<- sp.importance.fil |> 
  group_by(latitude, longitude, ProvId) |>
  summarize(sum = sum(R, na.rm=TRUE), .groups="keep") |> 
  ungroup() |> 
  right_join(grid, by = c("latitude", "longitude","ProvId")) |> 
  drop_na(ProvId) |> 
  select(longitude, latitude, sum)|> 
  mutate(sum=scales::rescale(sum))|> 
  drop_na()

# calculate the average rarity
dt1.fil <- sp.importance.fil |> 
  group_by(latitude, longitude, ProvId) |>
  summarize(sum = sum(R, na.rm = TRUE) / n(),
            .groups = "keep") |> 
  ungroup() |> 
  right_join(grid, by = c("latitude", "longitude", "ProvId")) |> 
  drop_na(ProvId) |>
  select(longitude, latitude, sum) |> 
  mutate(sum = scales::rescale(sum)) |> 
  drop_na()

# make a data frame that gets rarity for each lat/lon and also create a column for each ocean basin
diver_biom.fil<- ENSEMBLE_occ.wmean.tot |> 
  rename(SR =total) |> # number of species per lat/lon
  left_join(grid|>  select("longitude","latitude","ProvId"))|>
  left_join(dt.fil |>  rename(tot_rar=sum))|>
  left_join(dt1.fil |>  rename(avrg_rar=sum))|> 
  mutate(
    basin = case_when(
      ProvId %in% c(1) ~ "Arctic",
      ProvId %in% c(2, 3, 4, 5, 6, 11, 12) ~ "North Pacific",
      ProvId %in% c(7, 8, 9 , 10, 13) ~ "South Pacific",
      ProvId %in% c(14, 15, 16, 17, 18, 19, 20) ~ "Indian Ocean",
      ProvId %in% c(21, 22, 23, 24, 25, 26) ~ "North Atlantic",
      ProvId %in% c(27, 28, 29, 30) ~ "South Atlantic",
      ProvId %in% c(31, 32, 33) ~ "Southern Ocean",
    )) 

diver_biom.fil
```

# 2. Species Richness

SR were calculated for unrestricted range.
```{r  fig.height=5, fig.width=12, warning=FALSE}
# create a LDG of species richness graph
ldg_UR<- ENSEMBLE_occ.wmean.tot[, c(1,2,4)] |> filter(total>0) |>
  ltg()+
  geom_line(data=max.sp_lat , aes(x=latitude,y= count),color="grey70")+
  labs(x="Latitude (cell midpoint)",color="Species Richness",y="Species Richness")
# create species richness graph
SR_UR<- plot.map(ENSEMBLE_occ.wmean.tot[,c("longitude","latitude","total")] )+labs(color="Species Richness")
# arrange the plots on the same grid
sr_patch<- SR_UR + 
ldg_UR+ theme(legend.position = "none")+
  plot_layout(widths = c(3.0, 1))+
  #plot_annotation(tag_levels = 'A')+ 
  plot_layout(guides = 'collect')
sr_patch
```

# 3. Rarity Maps + LDG

## Unfiltered

```{r rarity, fig.height=5, fig.width=12, warning=FALSE}
# total range-size rarity
tot_rarity<- plot.map(dt )+labs(color="Total range-size rarity")+
    scale_colour_gradientn(colours =  my.colors,
    trans="sqrt",
    guide = guide_colourbar(
                          title.position = "left",  barwidth=0.5,barheight=10,ticks=TRUE,
                          nbin = 10))
# total range-size rarity LDG
ldg_tot_rarity<- dt  |> rename(total=sum) |> 
  ltg() +labs(x="Latitude (cell midpoint)",y="Total range-size rarity",color="Total range-size rarity")+
    scale_colour_gradientn(colours =  my.colors,
    trans="sqrt",
    guide = guide_colourbar(
                          title.position = "left",  barwidth=0.5,barheight=10,ticks=TRUE,
                          nbin = 10))
# arrange the plots on the same grid
tot_rarity_patch<- tot_rarity+ theme(legend.position="none") + ldg_tot_rarity+ 
  plot_layout(widths = c(3.0, 1.))+
  #plot_annotation(tag_levels = 'A')+ 
  plot_layout(guides = 'collect')
tot_rarity_patch

# Average range-size rarity
avrg_rarity<- plot.map(dt1)+labs(color="Average range-size rarity")+
    scale_colour_gradientn(colours =  my.colors,
    trans="sqrt",
    guide = guide_colourbar(
                          title.position = "left",  barwidth=0.5,barheight=10,ticks=TRUE,
                          nbin = 10))
# Average range-size LDG
ldg_avrg_rarity<- dt1  |> rename(total=sum) |> 
  ltg() +labs(x="Latitude (cell midpoint)",y="Average range-size rarity",color="Average range-size rarity")+      scale_colour_gradientn(colours =  my.colors,
    trans="sqrt",
    guide = guide_colourbar(
                          title.position = "left",  barwidth=0.5,barheight=10,ticks=TRUE,
                          nbin = 10))
# arrange the plots on the same grid
avrg_rarity_patch<-avrg_rarity+ 
  theme(legend.position="none") + 
  ldg_avrg_rarity+ 
  plot_layout(widths = c(3.0, 1.))+
  #plot_annotation(tag_levels = 'A')+ 
  plot_layout(guides = 'collect')
avrg_rarity_patch

```

For 2D maps:

- The Jenks optimization method, also called the Jenks natural breaks classification method, is a data clustering method designed to determine the best arrangement of values into different classes. This is done by seeking to minimize each class's average deviation from the class mean, while maximizing each class's deviation from the means of the other classes. In other words, the method seeks to reduce the variance within classes and maximize the variance between classes.[
- Classes can be applied with the style parameter using four approaches for calculating breaks: "quantile" (default), "equal", "fisher", and "jenks". The default "quantile" approach will create relatively equal “buckets” of data for mapping, with a break created at the median (50th percentile) for a two-by-two map or at the 33rd and 66th percentiles for a three-by-three map. For a four-by-four map, breaks are created at the 25th, 50th (median), and 75th percentiles. 


```{r 2D, fig.width=12}
# create map with 2D legend for SR and rarity
dt2D <- full_join(ENSEMBLE_occ.wmean.tot[, c(1, 2, 4)], dt, by = c("longitude", "latitude")) |> drop_na(sum, total)
SR_TotRar_2D <- plot.map2D(dt2D,
                           dim = 4,
                           xlab = "species richness",
                           ylab = "total rarity")

dt2D <- full_join(ENSEMBLE_occ.wmean.tot[, c(1, 2, 4)], dt1, by = c("longitude", "latitude")) |> drop_na(sum, total)
SR_AvrgRar_2D <- plot.map2D(dt2D,
                            dim = 4,
                            xlab = "species richness",
                            ylab = "average rarity")

# arrange the plots on the same grid
patch2D <- SR_TotRar_2D + SR_AvrgRar_2D
patch2D
```

### All plots combined

```{r FigureS8, fig.height=20, fig.width=15, warning=FALSE}
# combine all pots togather on teh same grid
FigureS8<-(tot_rarity_patch/
avrg_rarity_patch/
patch2D) +
  plot_layout(heights = c(1, 1, 1))+ 
  plot_annotation(tag_levels = 'A')& 
  theme(plot.tag = element_text("Times New Roman",face = "bold", size = 12,hjust = -3, vjust = 0))
FigureS8
ggsave("img/FigureS8.png", width=15, height=20, units="in", dpi=300)
```

## Filtered

Note: Removed 2 sp Pacific Gyre


```{r rarity-fil, fig.height=5, fig.width=12}
# total range-size rarity
tot_rarity.fil<- plot.map(dt.fil )+labs(color="Total range-size rarity")

ldg_tot_rarity.fil<- dt.fil  |> rename(total=sum) |> 
  ltg() +labs(x="Latitude (cell midpoint)",y="Total range-size rarity",color="Total range-size rarity")

tot_rarity_patch.fil<- tot_rarity.fil+ theme(legend.position="none") + ldg_tot_rarity.fil+ 
  plot_layout(widths = c(3.0, 1))+
  plot_layout(guides = 'collect')
tot_rarity_patch.fil

# Average range-size rarity
avrg_rarity.fil<- plot.map(dt1.fil)+labs(color="Average range-size rarity")

ldg_avrg_rarity.fil<- dt1.fil  |> rename(total=sum) |> 
  ltg() +labs(x="Latitude (cell midpoint)",y="Average range-size rarity",color="Average range-size rarity")

avrg_rarity_patch.fil<-avrg_rarity.fil+ theme(legend.position="none") + ldg_avrg_rarity.fil+ 
  plot_layout(widths = c(3.0, 1))+
  #plot_annotation(tag_levels = 'A')+ 
  plot_layout(guides = 'collect')
avrg_rarity_patch.fil

```


Filtered total range-size rarity map (rarity values >0.7 were removed)
```{r FigureS3, fig.height=5, fig.width=12}
# total range-size rarity
tot_rarity.filtered<- plot.map(dt.fil |> filter(sum<0.7) )+labs(color="Total range-size rarity")

ldg_tot_rarity.filtered<- dt.fil  |> rename(total=sum) |> filter(total<0.7)|> 
  ltg() +labs(x="Latitude (cell midpoint)",y="Total range-size rarity",color="Total range-size rarity")

FigureS3<- tot_rarity.filtered+ theme(legend.position="none") + ldg_tot_rarity.filtered+ 
  plot_layout(widths = c(3.0, 1))+
  #plot_annotation(tag_levels = 'A')+ 
  plot_layout(guides = 'collect')
FigureS3

ggsave("img/FigureS9.png",plot=FigureS3, width=12, height=5, units="in", dpi=300)
```

fro 2D legends:

- The Jenks optimization method, also called the Jenks natural breaks classification method, is a data clustering method designed to determine the best arrangement of values into different classes. This is done by seeking to minimize each class's average deviation from the class mean, while maximizing each class's deviation from the means of the other classes. In other words, the method seeks to reduce the variance within classes and maximize the variance between classes.[
- Classes can be applied with the style parameter using four approaches for calculating breaks: "quantile" (default), "equal", "fisher", and "jenks". The default "quantile" approach will create relatively equal “buckets” of data for mapping, with a break created at the median (50th percentile) for a two-by-two map or at the 33rd and 66th percentiles for a three-by-three map. For a four-by-four map, breaks are created at the 25th, 50th (median), and 75th percentiles. 

```{r 2D-fil, fig.height=5, fig.width=12, warning=FALSE}

dt2D.fil <- full_join(ENSEMBLE_occ.wmean.tot[, c(1,2,4)], dt.fil, by = c("longitude", "latitude")) |> drop_na(sum,total)
SR_TotRar_2D.fil <- plot.map2D(dt2D.fil, dim=4,  xlab = "species richness",ylab = "total rarity")

dt2D.fil <- full_join(ENSEMBLE_occ.wmean.tot[, c(1,2,4)],dt1.fil, by = c("longitude", "latitude")) |> drop_na(sum, total)
SR_AvrgRar_2D.fil<- plot.map2D(dt2D.fil, dim=4,  xlab = "species richness",ylab = "average rarity")

patch2D.fil <- SR_TotRar_2D.fil + SR_AvrgRar_2D.fil+ 
  plot_annotation(tag_levels = 'A')& 
  theme(plot.tag = element_text("Times New Roman",face = "bold", size = 12,hjust = -2, vjust = 1.5))
patch2D.fil
```

### All Plots together
```{r Figure1, fig.height=15, fig.width=15, warning=FALSE}
# arrange the plots on the same grid
Figure1<-(sr_patch /
tot_rarity_patch.fil /
avrg_rarity_patch.fil ) +
  plot_layout(heights = c(1, 1, 1))+ 
  plot_annotation(tag_levels = 'A')& 
  theme(plot.tag = element_text("Times New Roman",face = "bold", size = 12,hjust = -3, vjust = 0))
Figure1
ggsave("img/Figure1.png", width=30, height=42, units="cm", dpi=300)
```

```{r Figure2, fig.height=5, fig.width=15, warning=FALSE}
# arrange the plots on the same grid
Figure2<-SR_TotRar_2D.fil + SR_AvrgRar_2D.fil+ 
  plot_annotation(tag_levels = 'A')& 
  theme(plot.tag = element_text("Times New Roman",face = "bold", size = 12,hjust = -2, vjust = 1.5))
Figure2
ggsave("img/Figure2.png", width=15, height=5, units="in", dpi=300)
```

# 4. Metabolic Theory Plots

The strong role of temperature is consistent with the metabolic 
theory of ecology (12), but while this theory implies a single negative 
and linear relationship with a slope of −0.32 between logarithmic 
species richness and inverse thermal energy (22, 23),
```{r Figure3, fig.height=6, fig.width=6}
# data frame with temperature, species richness, and hemisphere information
data <-  diver_biom.fil |>  left_join(Temp, by = c("longitude", "latitude")) |>
  mutate(lat = ifelse(latitude < 0, "Southern Hemisphere", "Northern Hemisphere")) |>
  select(avrg_rar,
         Temperature = Temp,
         SR = SR,
         Hemisphere = lat,
         avrg_rar)

# calculate ln(SR) and inverse thermal energy in eV^-1
data$ln_SR <- log(data$SR)
data$ln_avrg_rar <- log(data$avrg_rar)
k <- 8.617333262145e-05 # Boltzmann constant in eV/K
data$inv_kT <- 1 / (k * (data$Temperature + 273.15))


text <- data |> drop_na(Temperature) |>
  mutate(areas = case_when(Temperature < 5 ~ "left",
                           Temperature < 19 ~ "middle",
                           TRUE ~ "right")) |> tabyl(areas) |> as.data.frame() |>
  mutate(
    x = c(1, 12, 23),
    x = 1 / (k * (x + 273.15)),
    y = 0.7,
    area = paste0(round(percent * 100, 0), "% area")
  )

# ggplot with geom_point
Figure3<-ggplot(data, aes(x = inv_kT, y = ln_SR, group = Hemisphere)) +
    geom_vline(xintercept = 1 / (k * (5 + 273.15)), color="grey")+
  geom_vline(xintercept = 1 / (k * (19 + 273.15)), color="grey")+
  geom_point(aes(color = Hemisphere),alpha=0.05, size=1, show.legend = F) +
    geom_abline(slope = 0.70, intercept=32.72, linetype = "dashed")+ # slope should be - -> but we reverse the scale
  scale_y_continuous(name = "ln(SR)") +
  labs(color = "Hemisphere") +
  theme_classic() +
  theme(legend.position = "right") +
   # add a secondary x axis with temperature in Celsius
  scale_x_reverse(sec.axis=sec_axis(trans=~ (1/.)/k-273.15, name="Temperature (°C)",
                                    breaks=pretty_breaks(n=5)))+
  scale_y_continuous(breaks=pretty_breaks(n=5))+
  labs(x=TeX(r'(\textbf{Inverse Thermal Energy $(eV^{-1})$})'))+
  scale_color_manual(values=c("#ffe169","#4893c6"))+
  ggnewscale::new_scale_color() +
  geom_smooth(method = "loess", span=0.5, se=FALSE, aes(color=Hemisphere))+
   scale_color_manual(values=c("#cc9c00","#054a91"))+
  labs(y='Species Richness (ln)')+
  geom_text(data = text, aes(x, y, label =area),inherit.aes = FALSE)+
  theme_Publication()+ 
  theme(legend.position = c(0.85, 0.335))
Figure3
ggsave("img/Figure3.png", width=6, height=6, units="in", dpi=300)
```


# 5. SR vsTemp

```{r T_SR}
T_SR.plt <- diver_biom.fil |> left_join(Temp, by = c("longitude", "latitude")) |>
  select(Temp,SR,avrg_rar) |> 
  drop_na() |> 
 arrange(avrg_rar) |> 
  ggplot(aes(x=Temp, y=SR, color=avrg_rar)) +
  geom_point(alpha=0.5)+
  scale_color_gradientn(colours = my.colors,
                        #trans="log10",
                        guide = guide_colourbar(title.position = "left",
                      barwidth=0.5,barheight=10,ticks=TRUE,
                          nbin = 10))+
 
  theme_Publication()+
  geom_smooth(color="black",method = "loess", se=FALSE)+
  theme(legend.position = "right",
            legend.key.size= unit(0.2, "cm"),
            legend.spacing = unit(0, "cm"),
            legend.title = element_text(face="bold",angle = 90),
        legend.title.align = 0.5)+
   labs(x="Temperature (°C)", y="Species Richness", color="Average range-size rarity")+
  scale_x_continuous(breaks = pretty_breaks(n=10))



Tsd_map <-diver_biom.fil |> left_join(MESO, by = c("longitude", "latitude")) |> 
  select(longitude, latitude, Tsd) |>
  plot.map()+
  labs(color="Temperature Std Dev")


Tsd_SR.plt <- diver_biom.fil |> left_join(MESO, by = c("longitude", "latitude"))|>
  rename(Temp=Tsd) |> 
  select(Temp,SR,Tmean) |> 
  drop_na() |> 
 #arrange(Tmean) |> 
  ggplot(aes(x=Tmean, y=SR, color=Temp)) +
  geom_point(alpha=0.5)+
  scale_color_gradientn(colours = my.colors,
                        #trans="log10",
                        guide = guide_colourbar(title.position = "left",
                      barwidth=0.5,barheight=10,ticks=TRUE,
                          nbin = 10))+
  theme_Publication()+
  geom_smooth(color="black",method = "loess", se=FALSE)+
  theme(legend.position = "right",
            legend.key.size= unit(0.2, "cm"),
            legend.spacing = unit(0, "cm"),
            legend.title = element_text(face="bold",angle = 90),
        legend.title.align = 0.5)+
   labs(x="Temperature (°C)", y="Species Richness", color="Temperature Std Dev")+
  scale_x_continuous(breaks = pretty_breaks(n=10))#+
 # scale_y_log10()

```

```{r FigureS4,fig.height=12, fig.width=15}
FigureS4<-(T_SR.plt | Tsd_SR.plt) / Tsd_map +
  plot_layout(heights = c(1, 1.5))+ 
  plot_annotation(tag_levels = 'A')& 
  theme(plot.tag = element_text("Times New Roman",face = "bold", size = 12,hjust = -3, vjust = 0))
FigureS4
ggsave("img/FigureS4.png", width=15, height=12, units="in", dpi=300)
```



# 6. Sampling effort

```{r FigureS12, fig.height=12, fig.width=15, warning=FALSE}
# Num of SR effort
effort2<-plot.map(effort.infer|>  select(longitude, latitude, num_sp_obs) |>  drop_na() |> arrange(num_sp_obs),alpha=0.5)+
  labs(color="Number of Obs Species (sqrt)")+
  scale_color_gradientn(colours = my.colors,
                        trans="sqrt",
                        guide = guide_colourbar(title.position = "left", barwidth = 0.5,barheight = 10,ticks = TRUE,nbin = 10)) 

### LDG Rarirty vs Effort 
bias.df<- left_join(effort.infer, diver_biom.fil) |>  
  select(-num_sp_pred, -effort,  -basin) |>  drop_na()

bias_lat<- bias.df |>  group_by(latitude) |> 
  summarize(mean_sp= mean(num_sp_obs, na.rm=TRUE),
            sd_sp= sd(num_sp_obs, na.rm=TRUE),
            mean_samples= mean(n_samples_obs, na.rm=TRUE),
            sd_ssamples= sd(n_samples_obs, na.rm=TRUE))

p1<- ldg_avrg_rarity.fil+
  theme(legend.position = "none")

p2<- ggplot(bias.df,aes(y=num_sp_obs, x=latitude))+ 
  geom_point(color="grey")+
  labs(x=NULL)+
  stat_summary( fun=mean, geom="line", colour="darkblue")+
  geom_smooth(method ="loess", se=TRUE)+
  scale_y_log10()+
  coord_flip()+
  theme_Publication()+
  labs(y= "Number of observed species")+
  theme(legend.title = element_text(colour="black", size=10, face="bold",angle = 90),
        legend.title.align = 0.5)+
  scale_x_continuous(breaks= c(-60,-30,0,30,60),
                              labels=c("60°S","30°S","0","30°N","60°N"),
                              limits=c(-80,90) )+ 
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank())

# plot_grid(plot_grid(p1, p2, ncol=2,labels = "AUTO"), plot_grid(effort2, labels = "C"), ncol=1)
FigureS12<-(p1 | p2) / effort2 +
  plot_layout(heights = c(1, 1.5))+ 
  plot_annotation(tag_levels = 'A')& 
  theme(plot.tag = element_text("Times New Roman",face = "bold", size = 12,hjust = -3, vjust = 0))
FigureS12
ggsave("img/FigureS12.png", width=15, height=12, units="in", dpi=300)
```

# 7. ENV Maps

```{r FigureS12, fig.height=10, fig.width=11}
load(here::here("rds/ENV.RData"))
#MESO
FigureS12<-cowplot::plot_grid(
plot.map(MESO |> select(longitude,latitude, Temp))+labs(color="Temp"),
plot.map(MESO |> select(longitude, latitude, Sal) |> filter(Sal>25)) + labs(color = "Sal") +
  scale_color_gradientn(
    colours = my.colors,
    trans="log10",
    guide = guide_colourbar(
      title.position = "left",
      barwidth = 0.5,
      barheight = 10,
      ticks = TRUE,
      nbin = 10
    )
  ) ,
plot.map(MESO |> select(longitude,latitude, O2))+labs(color="O2"),
plot.map(MESO |> select(longitude,latitude, NO3))+labs(color="NO3"),
plot.map(MESO |> select(longitude,latitude, MLD))+labs(color="MLD"),
plot.map(MESO |> select(longitude,latitude, Zeu))+labs(color="Zeu"),
plot.map(MESO |> select(longitude,latitude, NPP))+labs(color="NPP"), labels = "AUTO", ncol = 2
)
FigureS12
ggsave("img/FigureS12.png", width=29.7, height=25, units="cm", dpi=300)
```

# 8. Plots per Phylum

Create a combined Figures 1 and 2 - but for each phylum.

For 2D legend maps : The terminology for dim when using bi_class refers to how many categories each variable (e.g., species richness and rarity) is divided into. Here’s how to interpret it:

dim = 2: Each variable is divided into two categories (low and high). This results in a 2x2 grid (4 classes) when plotting the bivariate map.
dim = 3: Each variable is divided into three categories (low, medium, and high). This creates a 3x3 grid (9 classes) on the bivariate map.
dim = 4: Each variable is divided into four categories (very low, low, medium, and high). This leads to a 4x4 grid (16 classes) on the bivariate map.
When calling bi_class, these values indicate the granularity of the classification, with higher dim values resulting in finer differentiation between categories.

```{r FigureS1, fig.height=20, fig.width=15}
# For Figure S1 (Annelida with dim = 2)
plot_per_phylum("Annelida", 2)
ggsave("img/FigureS1.png", width=30, height=42, units="cm", dpi=300)

# For Figure S2 (Arthropoda with dim = 4)
plot_per_phylum("Arthropoda", 4)
ggsave("img/FigureS2.png", width=30, height=42, units="cm", dpi=300)

# For Figure S3 (Chaetognatha with dim = 2)
plot_per_phylum("Chaetognatha", 2)
ggsave("img/FigureS3.png", width=30, height=42, units="cm", dpi=300)

# For Figure S4 - Chordata (style = "equal", dim = 3)
plot_per_phylum("Chordata", 3, "img/FigureS4.png", style = "equal")

# For Figure S5 - Cnidaria (style = "quantile", dim = 4)
plot_per_phylum("Cnidaria", 4, "img/FigureS5.png", style = "quantile")

# For Figure S6 - Ctenophora (style = "equal", dim = 3)
plot_per_phylum("Ctenophora", 3, "img/FigureS6.png", style = "equal")

# For Figure S7 - Mollusca (style = "quantile", dim = 4)
plot_per_phylum("Mollusca", 4, "img/FigureS7.png", style = "quantile")
```

